Reparametrization of COM-Poisson Regression Models with Applications in the Analysis of Experimental Data

### 2018-01-29 article.Rnw
Correct spelling

To check the orthogonality property we compute the covariance matrix
between maximum likelihood estimators
$\hat{\bm{\theta}} = (\hat{\bm{\beta}}, \phi)$, obtained by observed
information matrix,
Cov$(\hat{\bm{\theta}}) = \mathcal{I}^{-1}(\bm{\theta})$.
\autoref{fig:ortho-plot} shows the covariance between regression and
dispersion parameter estimators for each simulation scenario, on
the correlation scale. The correlations are close to zero in all cases
suggesting the orthogonality property for the reparametrized
model. Interesting results in the first panel shows that
cov($\hat{\beta}_{22}, \hat{\phi}$) is not very close to zero (values
between $-0.4$ and $0.2$) for strong overdispersion ($\phi = -1.6$).

@@

<<ortho-surf, fig.height=5, fig.width=9, fig.cap="Deviance surfaces under original and proposed parametrization for simulated data sets of size 1000 an different dispersion parameters. The ellipses are confidence regions (90, 95 and 99\\%), dotted lines are the maximum likelihood estimates, and points are the real parameters used in the simulation.">>=

@@

To ilustrate the orthogonality, \autoref{fig:ortho-surf} the deviance
surfaces for four simulated data set of size 1000, $\mu=5$ and different
values of the dispersion parameters. The shapes of the deviance function
show that the proposed parametrization is better for both computation
and asymptotic (normal-based) inference. Furthermore, it is interesting
to note that the deviance function shape on strong overdispersion
($\phi=-1.6$) not as well behaved as the others. This is due to the
difficulty of the distribution in dealing with strong overdispersion in
low counts (see dispersion index plot in the
\autoref{fig:indexes-plot}). This also explains the results of
Cov$(\hat{\beta}_{22}, \phi)$ in the first panel of
\autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and associate
with low counts.

@@

<<desc-soy, fig.pos="!ht", fig.height=4, fig.width=8, fig.cap="Number of bean seeds per pot for each potassium dose and moisture level (left) and sample mean against sample variance of the five replicates for each experimental treatment (right). Solid lines are the smoothing curves on the left and the least of squares curve on the right.">>=

@@

<<pred-soy, fig.pos="!ht", fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of been seeds counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=

>>>>

To check the orthogonality property we compute the covariance matrix
between maximum likelihood estimators
$\hat{\bm{\theta}} = (\hat{\bm{\beta}}, \phi)$, obtained from the
observed information matrix,
Cov$(\hat{\bm{\theta}}) = \mathcal{I}^{-1}(\bm{\theta})$.
\autoref{fig:ortho-plot} shows the covariance between regression and
dispersion parameter estimators for each simulation scenario, on the
correlation scale. The correlations are close to zero in all cases
suggesting the orthogonality property for the reparametrized
model. Interestingly, results in the first panel show that
cov($\hat{\beta}_{22}, \hat{\phi}$) is not very close to zero (values
between $-0.4$ and $0.2$) for strong overdispersion ($\phi = -1.6$).

@@

<<ortho-surf, fig.height=5, fig.width=9, fig.cap="Deviance surfaces contour plots under original and proposed parametrization for four simulated data sets of size 1000 with different dispersion parameters. The ellipses are confidence regions (90, 95 and 99\\%), dotted lines are the maximum likelihood estimates, and points are the real parameters used in the simulation.">>=

@@

To illustrate the orthogonality, \autoref{fig:ortho-surf} displays
contour plots of the deviance surfaces for four simulated data set of
size 1000, $\mu=5$ and different values of the dispersion
parameters. The shapes of the deviance function show that the proposed
parametrization is better for both computation and asymptotic
(normal-based) inference. Furthermore, it is interesting to note that
the deviance function shape under strong overdispersion ($\phi=-1.6$) is
not as well behaved as the others; this is due to the difficulty of the
distribution in dealing with strong overdispersion in low counts (see
dispersion index plot in the \autoref{fig:indexes-plot}). This also
explains the results of Cov$(\hat{\beta}_{22}, \phi)$ in the first panel
of \autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and
associated with low counts.

@@

<<desc-soy, fig.height=4, fig.width=8, fig.cap="Number of bean seeds per pot for each potassium dose and moisture level (left) and sample mean against sample variance of the five replicates for each experimental treatment (right). Solid lines are the smoothing curves on the left and the least of squares curve on the right.">>=

@@

<<pred-soy, fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of been seeds counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=
